Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
8 pages!

Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We interleave wake cycles with each of the sleep cycles;
(2) on the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a DSL has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
(3) Because the frontiers can grow very large,
we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(4) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ -- we found that $K = 2$ sufficed.
(5) For added robustness, we enumerate programs from both the generative model and the recognition model.

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks along with input/output examples (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot find any programs that
work at all for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14,muggleton2015meta}.

>>>>

Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We interleave wake cycles with each of the sleep cycles;
(2) on the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a DSL has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
%% (3) Because the frontiers can grow very large,
%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
(4) For added robustness, we enumerate programs from both the generative model and the recognition model.

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and recover many of the higher-level building blocks that have made these
other system successful.
Tbl.~\ref{textPrimitives} (top) lists the primitive functions provided to \systemEnding.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide list/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}). %, split  75/25 train/test.
At first, \system cannot find any programs that
work at all for most of the tasks.
After three wake/sleep iterations,
it assembles a DSL (Tbl.~\ref{textPrimitives}, bottom) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, and predicate invention in ILP~\cite{DBLP:conf/ecai/LinDETM14}.
